Comparison of Terrain Following and Cut Cell Grids with a Non-Hydrostatic Model

### 2015-02-10 fig-advection-error.tex
add subfigure labels, improve captioning

\documentclass{article}
\usepackage[paperwidth=3in,paperheight=3.0in,margin=0in]{geometry}
\usepackage{mathptmx}%{times}
\usepackage{graphicx}
\usepackage{siunitx}

\setlength{\parindent}{0pt}
\begin{document}
\hspace{0.15in}\includegraphics[width=2.8in,height=1.32in]{img/schaer-btf-4thorder-error.png}

\footnotesize
\input{../advectTracer-horizontal-btf-cubicUpwindCPCFit-divFree/error-contour-x-only}
\end{document}

>>>>

\documentclass{article}
\usepackage[paperwidth=3in,paperheight=3.0in,margin=0in]{geometry}
\usepackage{mathptmx}%{times}
\usepackage{graphicx}
\usepackage{siunitx}
\usepackage[absolute,overlay]{textpos}

\setlength{\parindent}{0pt}
\begin{document}
\begin{textblock}{1}(2.5,0.4)
\normalsize
(a)
\end{textblock}
\begin{textblock}{1}(2.5,8.5)
\normalsize
(b)
\end{textblock}
\hspace{0.35in}\includegraphics[width=2.6in,height=1.32in]{img/schaer-btf-4thorder-error.png}

\footnotesize
\input{../advectTracer-horizontal-btf-cubicUpwindCPCFit-divFree/error-contour}
\end{document}

### 2015-02-10 msc-publication.tex
add subfigure labels, improve captioning

It is well-known that TF coordinates perform badly in the presence of steep orography: as model resolution increases, steeper gradients lead to larger truncation errors in calculating the horizontal pressure gradientwhich result in spurious winds \citep{dempsey-davis1998,steppeler2002}.

Much work has been done to reduce error associated with TF coordinates: firstly by smoothing the effects of terrain with height \citep{schaer2002,klemp2011} and, secondly, by improving the accuracy in calculating the horizontal pressure gradient itself \citep{klemp2011} \TODO{and zaengl2012?}.

@@

\section{Numerical method}

@@

\begin{figure}
	\centering
	\includegraphics{../msc-publication-fig-advection-tracer/fig-advection-tracer.pdf}
%
	\caption{Horizontally advected tracer contours at $t = \SI{0}{\second}$, \SI{5000}{\second} and \SI{10000}{\second}.  \TODO{identify subfigures}  Contour intervals are every 0.1.}
	\label{fig:advection-tracer}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics{../msc-publication-fig-advection-error/fig-advection-error.pdf}
	%
	\caption{Errors in horizontal tracer advection at $t = \SI{10000}{\second}$.  \TODO{identify subfigures}  Contour intervals are every 0.01 with negative contours denoted by dashed lines.}
	\label{fig:advection-error}
\end{figure}

@@

\begin{figure*}
	\centering
	\includegraphics{../msc-publication-fig-resting/fig-resting.pdf}
%
	\caption{Maximum spurious vertical velocity $w$ in the resting atmosphere test \TODO{identify subfigures}.  Note that vertical scales differ.}
	\label{fig:resting}
\end{figure*}

@@

\begin{figure*}
	\centering
	\includegraphics{../msc-publication-fig-gravityWaves-theta/fig-gravityWaves-theta.pdf}
%
	\caption{Anomalies in potential temperature in the gravity waves test after 5 hours.  \TODO{identify subfigures}}
	\label{fig:gw-theta}
\end{figure*}

\begin{figure}
	\centering
	\includegraphics{../msc-publication-fig-gravityWaves-sampleLine/fig-gravityWaves-sampleLine.pdf}
%
	\caption{Vertical profiles of the Exner function of pressure, $\Pi$, and potential temperature, $\theta$, in the gravity waves test.  The computational mode is manifested as a zig-zag in potential temperature on the SnapCol grid.  Results on the SLEVE grid (not shown) are qualitatively the same as those on the BTF grid.}
	\label{fig:gw-exner-theta}
\end{figure}

>>>>

It is well-known that TF coordinates perform badly in the presence of steep orography: as model resolution increases, steeper gradients lead to larger truncation errors in calculating the horizontal pressure gradient which result in spurious winds \citep{dempsey-davis1998,steppeler2002}.

Much work has been done to reduce error associated with TF coordinates: firstly by smoothing the effects of terrain with height \citep{schaer2002,klemp2011} and, secondly, by improving the accuracy in calculating the horizontal pressure gradient itself \citep{klemp2011,zaengl2012}.

@@

\TODO{cut cell method}

\section{Numerical method}
We use a finite volume discretisation of the fully-compressible Euler equations described by \citet{weller-shahrokhi2014}.  The model has a curl-free pressure gradient formulation, an upwind-biased cubic advection scheme, and uses a Lorenz staggering of thermodynamic variables.

@@

\begin{figure}
	\centering
	\includegraphics{../msc-publication-fig-advection-tracer/fig-advection-tracer.pdf}
%
	\caption{Horizontally advected tracer contours at $t = \SI{0}{\second}$, \SI{5000}{\second} and \SI{10000}{\second} using (a) the fourth-order centred difference scheme with sigma coordinates from \citet{schaer2002}, and the upwind-biased cubic scheme on (b) BTF grid, (c) cut cell grid.  Contour intervals are every 0.1.}
	\label{fig:advection-tracer}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics{../msc-publication-fig-advection-error/fig-advection-error.pdf}
	%
	\caption{Errors in horizontal tracer advection at $t = \SI{10000}{\second}$ using (a) the fourth-order centred difference scheme with sigma coordinates from \citet{schaer2002}, (b) the upwind-biased cubic scheme on a BTF grid.  Contour intervals are every 0.01 with negative contours denoted by dashed lines.}
	\label{fig:advection-error}
\end{figure}

@@

\begin{figure*}
	\centering
	\includegraphics{../msc-publication-fig-resting/fig-resting.pdf}
%
	\caption{Maximum spurious vertical velocity $w$ in the resting atmosphere test with results on (a) BTF, SLEVE, Hybrid Terrain Following (HTF) and Smoothed Terrain Following (STF) coordinates from \citet{klemp2011}, (b) BTF, SLEVE, cut cell and regular grids using the model from \citet{weller-shahrokhi2014}.  Note that vertical scales differ.}
	\label{fig:resting}
\end{figure*}

@@

\begin{figure*}
	\centering
	\includegraphics{../msc-publication-fig-gravityWaves-theta/fig-gravityWaves-theta.pdf}
%
	\caption{Anomalies in potential temperature in the gravity waves test after 5 hours.  The central domain in the lowest \SI{12}{\kilo\meter} is shown on (a) the BTF grid, and (c) the cut cell grid.  The four lowest layers of each grid are shown for (b) BTF, and (d) cut cell grids, using a narrower potential temperature scale.}
	\label{fig:gw-theta}
\end{figure*}

\begin{figure}
	\centering
	\includegraphics{../msc-publication-fig-gravityWaves-sampleLine/fig-gravityWaves-sampleLine.pdf}
%
	\caption{Vertical profiles of the Exner function of pressure, $\Pi$, and potential temperature, $\theta$, in the gravity waves test.  The computational mode is manifested as a zig-zag in potential temperature on the cut cell grid.  Results on the SLEVE grid (not shown) are qualitatively the same as those on the BTF grid.}
	\label{fig:gw-exner-theta}
\end{figure}
